---
title: "Analyses for Bachelor Thesis"
author: "Victoria Stefaanou"
date: '2024-05-21'
output:
  pdf_document: default
  html_document: default
---

Load data
```{r}
all_data_emo <- read.csv("/Users/victoriastefanou/Downloads/all_data_emo.csv")
all_data <- read.csv("/Users/victoriastefanou/Downloads/all_data.csv")
neg_all <- read.csv("/Users/victoriastefanou/Downloads/neg_all.csv")
pos_all <- read.csv("/Users/victoriastefanou/Downloads/pos_all.csv")
av_scores_final <- read.csv("/Users/victoriastefanou/Downloads/av_scores_final.csv")
av_mw_score <- read.csv("/Users/victoriastefanou/Downloads/av_mw_score.csv")
neg_desc <- read.csv("/Users/victoriastefanou/Downloads/neg_desc.csv")
pos_desc <- read.csv("/Users/victoriastefanou/Downloads/pos_desc.csv")
```

MW percentage
```{r}
library(ggplot2)

# Overall MW
all_data_emo_noNA <- all_data_emo[!is.na(all_data_emo$MW), ] # remove rows where MW == NA

# percentage of MW occurrences overall
total_count_mw_all <- nrow(all_data_emo_noNA)
mw1_count <- sum(all_data_emo_noNA$MW == 1, na.rm = TRUE)
percentage_mw1 <- (mw1_count / total_count_mw_all) * 100

# percentage of MW neu all
total_count_mw_neu <- sum(all_data_emo_noNA$Condition =='neu')
mw_neu <- sum(all_data_emo_noNA$MW[all_data_emo_noNA$Condition == 'neu'] == 1, na.rm = TRUE)
percentage_mw_neu <- (mw_neu / total_count_mw_neu) * 100

# percentage of MW emo all
total_count_mw_emo <- sum(all_data_emo_noNA$Condition =='emo')
mw_emo <- sum(all_data_emo_noNA$MW[all_data_emo_noNA$Condition == 'emo'] == 1, na.rm = TRUE)
percentage_mw_emo <- (mw_emo / total_count_mw_emo) * 100

# Negative condition MW
# percentage of MW neu (neg condition)
total_count_mw_neu_N <- sum(neg_all$Condition =='neu')
mw_neu_N <- sum(neg_all$MW[neg_all$Condition == 'neu'] == 1, na.rm = TRUE)
percentage_mw_neu_N <- (mw_neu_N / total_count_mw_neu_N) * 100

# percentage of MW neg (neg condition)
total_count_mw_neg_N <- sum(neg_all$Condition =='neg')
mw_neg_N <- sum(neg_all$MW[neg_all$Condition == 'neg'] == 1, na.rm = TRUE)
percentage_mw_neg_N <- (mw_neg_N / total_count_mw_neg_N) * 100

# Positive condition MW
# percentage of MW neu (pos condition)
total_count_mw_neu_P <- sum(pos_all$Condition =='neu')
mw_neu_P <- sum(pos_all$MW[pos_all$Condition == 'neu'] == 1, na.rm = TRUE)
percentage_mw_neu_P <- (mw_neu_P / total_count_mw_neu_P) * 100

# percentage of MW pos (pos condition)
total_count_mw_pos_P <- sum(pos_all$Condition =='pos')
mw_pos_P <- sum(pos_all$MW[pos_all$Condition == 'pos'] == 1, na.rm = TRUE)
percentage_mw_pos_P <- (mw_pos_P / total_count_mw_pos_P) * 100

# plot
par(mar = c(5, 8, 4, 2) + 0.1)
bp_mw <- barplot(
  c(percentage_mw1, percentage_mw_neu, percentage_mw_emo, percentage_mw_neu_N, percentage_mw_neg_N, percentage_mw_neu_P, percentage_mw_pos_P),
  names.arg = c("total", "neu/all", "emo/all", "neu/N", "neg/N", "neu/P", "pos/P"),
  xlab = "Condition",
  ylab = "% MW",
  col = c("lightskyblue1", "lightskyblue3", "lightskyblue4", "mistyrose2", "mediumpurple1", "mistyrose3", "mediumpurple3"),
  main = "Percentage MW",
  width = 3,
  xlim = c(0, 20),
  ylim = c(0, 60),
  space = 0,
)
text(
  x = bp_mw, 
  y = c(percentage_mw1, percentage_mw_neu, percentage_mw_emo, percentage_mw_neu_N, percentage_mw_neg_N, percentage_mw_neu_P, percentage_mw_pos_P), 
  labels = c(round(percentage_mw1, 2), round(percentage_mw_neu, 2), round(percentage_mw_emo, 2), round(percentage_mw_neu_N, 2), round(percentage_mw_neg_N, 2), round(percentage_mw_neu_P, 2), round(percentage_mw_pos_P, 2)), 
  pos = 3, 
  col = "black"
)

print(percentage_mw1)     # 43% MW in total
print(percentage_mw_neu)  # 45% MW in neu all
print(percentage_mw_emo)  # 41% MW in emo all

print(percentage_mw_neu_N) # 40% MW in neu (neg)
print(percentage_mw_neg_N) # 31% MW in neg (neg)

print(percentage_mw_neu_P) # 50% MW in neu (pos)
print(percentage_mw_pos_P) # 53% MW in pos (pos)
```

MW + PC percentage
```{r}
only_pc <- all_data_emo[all_data_emo$MW == 1, ]

# Overall PC
# percentage of MW + PC overall
only_pc$PC[is.na(only_pc$PC)] <- 0
total_count_pc <- nrow(only_pc)
pc <- sum(only_pc$PC== 1)
percentage_pc <- (pc / total_count_pc) * 100

# percentage of MW neu + PC
total_count_pc_neu <- sum(all_data_emo$MW[all_data_emo$Condition == 'neu'] == 1, na.rm = TRUE)
pc_neu <- sum(all_data_emo$PC[all_data_emo$Condition == 'neu'] == 1, na.rm = TRUE)
percentage_pc_neu <- (pc_neu / total_count_pc_neu) * 100

# percentage of MW emo + PC
total_count_pc_emo <- sum(all_data_emo$MW[all_data_emo$Condition == 'emo'] == 1, na.rm = TRUE)
pc_emo <- sum(all_data_emo$PC[all_data_emo$Condition == 'emo'] == 1, na.rm = TRUE)
percentage_pc_emo <- (pc_emo / total_count_pc_emo) * 100

# Negative Condition PC
# percentage of MW + PC neu (neg condition)
total_count_pc_neu_N <- sum(neg_all$MW[neg_all$Condition == 'neu'] == 1, na.rm = TRUE)
pc_neu_N <- sum(neg_all$PC[neg_all$Condition == 'neu'] == 1, na.rm = TRUE)
percentage_pc_neu_N <- (pc_neu_N / total_count_pc_neu_N) * 100

# percentage of MW + PC neg (neg condition)
total_count_pc_neg_N <- sum(neg_all$MW[neg_all$Condition == 'neg'] == 1, na.rm = TRUE)
pc_neg_N <- sum(neg_all$PC[neg_all$Condition == 'neg'] == 1, na.rm = TRUE)
percentage_pc_neg_N <- (pc_neg_N / total_count_pc_neg_N) * 100

# Positive Condition PC
# percentage of MW + PC neu (pos condition)
total_count_pc_neu_P <- sum(pos_all$MW[pos_all$Condition == 'neu'] == 1, na.rm = TRUE)
pc_neu_P <- sum(pos_all$PC[pos_all$Condition == 'neu'] == 1, na.rm = TRUE)
percentage_pc_neu_P <- (pc_neu_P / total_count_pc_neu_P) * 100

# percentage of MW + PC pos (pos condition)
total_count_pc_pos_P <- sum(pos_all$MW[pos_all$Condition == 'pos'] == 1, na.rm = TRUE)
pc_pos_P <- sum(pos_all$PC[pos_all$Condition == 'pos'] == 1, na.rm = TRUE)
percentage_pc_pos_P <- (pc_pos_P / total_count_pc_pos_P) * 100

par(mar = c(5, 8, 4, 1) + 0.1)

bp_pc <- barplot(
  c(percentage_pc, percentage_pc_neu, percentage_pc_emo, percentage_pc_neu_N, percentage_pc_neg_N, percentage_pc_neu_P, percentage_pc_pos_P),
  names.arg = c("total", "neu/all", "emo/all", "neu/N", "neg/N", "neu/P", "pos/P"),
  xlab = "Condition",
  ylab = "% MW+PC",
  col = c("lightskyblue1", "lightskyblue3", "lightskyblue4", "ivory2", "indianred", "ivory3", "brown"),
  main = "Percentage MW+PC",
  width = 3,
  xlim = c(0, 20),
  ylim = c(0, 65),
  space = 0,
)

text(
  x = bp_pc, 
  y = c(percentage_pc, percentage_pc_neu, percentage_pc_emo, percentage_pc_neu_N, percentage_pc_neg_N, percentage_pc_neu_P, percentage_pc_pos_P), 
  labels = c(round(percentage_pc, 2), round(percentage_pc_neu, 2), round(percentage_pc_emo, 2), round(percentage_pc_neu_N, 2), round(percentage_pc_neg_N, 2), round(percentage_pc_neu_P, 2), round(percentage_pc_pos_P, 2)), 
  pos = 3, 
  col = "black"
)

print(percentage_pc) # % MW+PC overall

print(percentage_pc_neu) # 34% MW + PC in neutral condition
print(percentage_pc_emo) # 47% MW + PC in emotional condition

print(percentage_pc_neu_N) # 33% MW + PC neu in neg condition
print(percentage_pc_neg_N) # 57% MW + PC neg in neg condition

print(percentage_pc_neu_P) # 35% MW + PC neu in pos condition
print(percentage_pc_pos_P) # 40% MW + PC pos in pos condition

# rows 150 (P34, Probe 10) & 301 (P33, Probe 1) have MW == 1 and PC == NA
```

Number of Fixations Plot
```{r}
# Number of Fixations 20s prior MW vs. No MW

# average number of fixations 20s prior to MW
numfix_MW <- mean(all_data_emo$NumFix[all_data_emo$Condition == "neu" & all_data_emo$MW == 1], na.rm = TRUE)
numfix_MW_emo <- mean(all_data_emo$NumFix[all_data_emo$Condition == "emo" & all_data_emo$MW == 1], na.rm = TRUE)

# average number of fixations 20s prior to NO MW
numfix_noMW <- mean(all_data_emo$NumFix[all_data_emo$Condition == "neu" & all_data_emo$MW == 0], na.rm = TRUE)
numfix_noMW_emo <- mean(all_data_emo$NumFix[all_data_emo$Condition == "emo" & all_data_emo$MW == 0], na.rm = TRUE)

bp <- barplot(
  c(numfix_noMW, numfix_MW, numfix_noMW_emo, numfix_MW_emo),
  names.arg = c("No MW (neu/all)", "MW (neu/all)", "No MW (emo/all)", "MW (emo/all)"),
  xlab = "Condition",
  ylab = "Average Number of Fixations",
  col = c("azure3", "lightpink3", "azure4", "lightpink4"),
  main = "Average Number of Fixations 20s Prior to MW probe",
  width = c(0.4, 0.4, 0.4, 0.4),
  ylim = c(0, 75),
  space = 0,
)
text(
  x = bp, 
  y = c(numfix_noMW, numfix_MW, numfix_noMW_emo, numfix_MW_emo), 
  labels = c(round(numfix_noMW, 2), round(numfix_MW, 2), round(numfix_noMW_emo, 2), round(numfix_MW_emo, 2)), 
  pos = 3, 
  col = "black"
)
```

```{r}
print(numfix_noMW)
print(numfix_MW)
print(numfix_noMW_emo)
print(numfix_MW_emo)
```

Scatterplot FixDur & NumFix
```{r}
# plot number of fixations and fixation duration based on condition + MW/ no MW - to check for outliers

# scatterplot numfix neu
subset_neu_MW <- all_data_emo[all_data_emo$Condition == "neu" & all_data_emo$MW == 1, ]
subset_neu_noMW <- all_data_emo[all_data_emo$Condition == "neu" & all_data_emo$MW == 0, ]

plot(subset_neu_MW$NumFix, 
     xlab = "NumFix",
     ylab = "",
     main = "NumFix Neu MW",
     col = "blue",  
     pch = 16       
)

plot(subset_neu_noMW$NumFix, 
     xlab = "NumFix",
     ylab = "",
     main = "NumFix Neu NO MW",
     col = "orange",  
     pch = 16       
)

# scatterplot fixdur neu
plot(subset_neu_MW$FixDur, 
     xlab = "FixDur",
     ylab = "",
     main = "FixDur Neu MW",
     col = "lightblue",  
     pch = 16       
)

plot(subset_neu_noMW$FixDur, 
     xlab = "FixDur",
     ylab = "",
     main = "FixDur Neu NO MW",
     col = "springgreen4",  
     pch = 16       
)
```

Fixation Duration Plot
```{r}
# Fixation Duration 20s prior MW vs. No MW

# average fixation duration 20s prior to MW
fixdur_MW <- mean(all_data_emo$FixDur[all_data_emo$Condition == "neu" & all_data_emo$MW == 1], na.rm = TRUE)
fixdur_MW_emo <- mean(all_data_emo$FixDur[all_data_emo$Condition == "emo" & all_data_emo$MW == 1], na.rm = TRUE)

# average fixation duration 20s prior to NO MW
fixdur_noMW <- mean(all_data_emo$FixDur[all_data_emo$Condition == "neu" & all_data_emo$MW == 0], na.rm = TRUE)
fixdur_noMW_emo <- mean(all_data_emo$FixDur[all_data_emo$Condition == "emo" & all_data_emo$MW == 0], na.rm = TRUE)

bp_2 <- barplot(
  c(fixdur_noMW, fixdur_MW, fixdur_noMW_emo, fixdur_MW_emo),
  names.arg = c("No MW (neu/all)", "MW (neu/all)", "No MW (emo/all)", "MW (emo/all)"),
  xlab = "Condition",
  ylab = "Average Durations of Fixations",
  col = c("honeydew3", "palegreen3", "honeydew4", "palegreen4"),
  main = "Average Durations of Fixations 20s Prior to MW probe",
  width = c(0.5, 0.5, 0.5, 0.5),
  ylim = c(0, 210),
  space = 0
)
text(
  x = bp_2, 
  y = c(fixdur_noMW, fixdur_MW, fixdur_noMW_emo, fixdur_MW_emo), 
  labels = c(round(fixdur_noMW, 2), round(fixdur_MW, 2), round(fixdur_noMW_emo, 2), round(fixdur_MW_emo, 2)), 
  pos = 3, 
  col = "black"
)
```

Two Samples t-test - Flesch Reading-Ease, text difficulty, emotional valence
```{r}
# Flesch Reading-Ease
flesch_neg <- c(71, 80)
flesch_pos <- c(71, 80)
flesch_t_test <- t.test(flesch_neg, flesch_pos)
print(flesch_t_test)

# perceived text difficulty
dif_t_test <- t.test(pos_desc$Q5_1, neg_desc$Q5_1)
print(dif_t_test)

# perceived emotional valence
neg_desc$Q6[neg_desc$Q6 == 2] <- 1
val_t_test <- t.test(pos_desc$Q6, neg_desc$Q6)
print(val_t_test)
```

GLMM MW rates - neu vs. emo
```{r}
# Linear Mixed Effects Model
# use glmer() to fit a function when DV is binary (MW)
library(lme4)
library(car)

# MW rates - all (Neutral vs. Emotional)
MW_emo <- glmer(MW ~ Condition + (1|Participant), data=all_data_emo, family = binomial)
summary(MW_emo)
```

GLMM MW - negative condition
```{r}
# MW rates - Negative Condition
MW_neg <- glmer(MW ~ Condition + (1|Participant), data=neg_all, family = binomial)
summary(MW_neg)

# p = 0.20704 (not significant)
```

GLMM MW - positive condition
```{r}
# MW rates - Positive Condition
MW_pos <- glmer(MW ~ Condition + (1|Participant), data=pos_all, family = binomial)
summary(MW_pos)

# p = 0.687 (not significant)
```

GLMM MW - neg vs. pos
```{r}
# MW rates - Emotional Condition (Positive vs Negative)
subset_data <- all_data[all_data$Condition %in% c('neg', 'pos'), ]
MW_pos_neg <- glmer(MW ~ Condition + (1|Participant), data=subset_data, family = binomial)
summary(MW_pos_neg)
```

GLMM MW+PC - neu vs. emo
```{r}
# MW+PC rates - all (Neutral vs Emotional)
MW_PC_emo <- glmer(PC ~ Condition + (1|Participant), data=only_pc, family = binomial)
summary(MW_PC_emo)
```

GLMM MW+PC - negative condition
```{r}
# MW+PC - Negative condition

# subset data: neg + MW == 1
neg_pc <- neg_all[neg_all$MW == 1, ]

PC_neg <- glmer(PC ~ Condition + (1|Participant), data=neg_pc, family = binomial, na.action = "na.exclude")
summary(PC_neg)
```

GLMM MW+PC - positive condition
```{r}
# MW+PC - Positive conditionn

# subset data: neg + MW == 1
pos_pc <- pos_all[pos_all$MW == 1, ]

PC_pos <- glmer(PC ~ Condition + (1|Participant), data=pos_pc, family = binomial, na.action = "na.exclude")
summary(PC_pos)
```

GLMM MW+PC - neg vs. pos
```{r}
# MW+PC - Emotional Condition (Positive vs Negative)
# subset data: neg/pos + MW == 1
neg_pos_pc <- subset_data[subset_data$MW == 1, ]

PC_neg_pos <- glmer(PC ~ Condition + (1|Participant), data=neg_pos_pc, family = binomial, na.action = "na.exclude")
summary(PC_neg_pos)
```

LMM MW - NumFix
```{r}
# MW rates - number of fixations & condition (neu vs. emo)
MW_numfix2 <- lmer(NumFix ~ MW + Condition + MW:Condition + (1|Participant), data = all_data_emo)
Anova(MW_numfix2, type = "II")
```

LMM MW - FixDur
```{r}
# MW rates - fixation duration & condition (neu vs. emo)
#MW_fixdur <- glmer(MW ~ FixDur + Condition + + FixDur:Condition + (1|Participant), data=all_data_emo, family = binomial)
#summary(MW_fixdur)

MW_fixdur2 <- lmer(FixDur ~ MW + Condition + MW:Condition + (1|Participant), data = all_data_emo)
#summary(MW_fixdur2)
Anova(MW_fixdur2, type = "II")
```

Paired t-test - Comprehension
```{r}
# MW rates - Comprehension score & conditions

# subsets per condition
av_neg <- av_scores_final[1:36, ]
av_pos <- av_scores_final[37:nrow(av_scores_final), ]

# MW rates - comprehension score in negative condition
t_test_neg <- t.test(Comp ~ Condition, data = av_neg, paired = TRUE)
print(t_test_neg)

# MW rates - comprehension score in positive condition
t_test_pos <- t.test(Comp ~ Condition, data = av_pos, paired = TRUE)
print(t_test_pos)

```

Two Samples t-test - Comprehension neg/pos
```{r}
# MW rates - comprehension score in neg vs pos condition

# subset only neg & pos condition
av_emo <- av_scores_final[av_scores_final$Condition %in% c('neg', 'pos'), ]

t_test_emo <- t.test(Comp ~ Condition, data = av_emo, paired = FALSE)
print(t_test_emo)
```

Boxplot - comprehension negative condition
```{r}
ggplot(av_neg, aes(x = Condition, y = Comp)) +
  geom_boxplot() +  
  labs(x = "Condition", y = "Comp") +  
  ggtitle("Boxplot of Comp by Condition")
```

Boxplot - comprehension positive condition
```{r}
ggplot(av_pos, aes(x = Condition, y = Comp)) +
  geom_boxplot() +  
  labs(x = "Condition", y = "Comp") +  
  ggtitle("Boxplot of Comp by Condition")
```

Correlation MW - Comp score
```{r}
correlation <- cor.test(av_mw_score$MW, av_mw_score$Comp)
correlation
```